### 2020-03-05 CalNotes.tex
Partial proof CLT

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenized equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenized equation, rather than to those of the homogenized one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}.

@@

Central limit theorem text.
\begin{theorem}\label{thm:CLT} Central limit theorem statement:
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \sqrt{T}\left(\widehat A^\epl_k(T) - A\right) = \Lambda, \quad \text{in law,}
	\end{equation}
	where $\Lambda \sim \mathcal N(0, \Gamma)$ and 
	\begin{equation}
		\Gamma = 2\sigma \E^{\rho^0}[V'(Z) \otimes V'(X)]^{-1}\E^{\rho^0}[V'(Z) \otimes V'(Z)]\E^{\rho^0}[V'(X) \otimes V'(Z)]^{-1} \corr{+ \dots}.
	\end{equation}
\end{theorem}

\begin{proof} Let us introduce the notation
	\begin{equation}
		\mathcal M_\epl \defeq \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ], \quad \mathcal M_0 \defeq \E^{\rho^0} [ V'(Z) \otimes V'(X) ],
	\end{equation}
	and the notation
	\begin{equation}
		\mathfrak p_\epl \defeq \E^{\rho^\epl} \left [ \frac{1}{\epl} p' \left ( \frac{X^\epl}{\epl} \right ) V'(Z^\epl) \right ]\
	\end{equation}
	and let us remark that from the proof of Theorem \ref{thm:mainTheorem} one can deduce the equalities
	\begin{equation}\label{eq:CTL_A}
		A = \frac1\delta \mathcal M_0^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)],
	\end{equation}
	and
	\begin{equation}\label{eq:CTL_alpha}
		\alpha = \mathcal M_\epl^{-1} \left(\frac1\delta \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] - \mathfrak p_\epl \right).
	\end{equation}
	Let us furthermore remark that the decomposition \eqref{eq:alphaDecomposition} yields
	\begin{equation}
		\sqrt{T}\left(\widehat A^\epl_k(T) - A\right) = \sqrt{T} \left(\alpha - A + I_1^\epl - I_2^\epl\right).
	\end{equation}
	Replacing the expression for $A$ and $\alpha$ given in \eqref{eq:CTL_A} and \eqref{eq:CTL_alpha}, we get
	\begin{equation}
	\begin{aligned}
	\sqrt{T}\left(\widehat A^\epl_k(T) - A\right) &= \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl \right) \\
	&+  \frac{\sqrt{T}}\delta \left( \mathcal M_\epl^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] - \mathcal M_0^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)] \right) \\
	&- \sqrt{T} I_2^\epl(T).
	\end{aligned}
	\end{equation}
	\corr{Third term is OK:} We rewrite the term involving $I_2^\epl(T)$ as
	\begin{equation}
	\sqrt{T} I_2^\epl(T) = \frac{\sqrt{2\sigma}}{\sqrt{T}} \widetilde M^{-1} Q^\epl_T,
	\end{equation}
	where, since $Z_t^\epl$ is adapted with respect to the natural filtration $\mathcal F_t$ of the Wiener process $W \defeq \left(W_t, t \geq 0\right)$, the quantity
	\begin{equation}
	Q^\epl_T \defeq \int_0^T V'(Z^\epl_t) \dd W_t,
	\end{equation}
	is a martingale whose quadratic variation is given by
	\begin{equation}
	\langle Q^\epl \rangle_T = \int_0^T V'(Z^\epl_t) \otimes V'(Z^\epl_t) \dd t. 
	\end{equation}
	Since the ergodic theorem guarantees that 
	\begin{equation}
	\lim_{T\to\infty}\frac{\langle Q^\epl \rangle_T}{T} = \E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)], \quad \text{in } L^1(\mu^\epl),
	\end{equation}
	the martingale central limit theorem gives 
	\begin{equation}
	\lim_{T\to\infty}\frac{1}{\sqrt{T}} Q^\epl_T = \Xi^\epl, \quad \text{in law},
	\end{equation}
	where $\Xi^\epl \sim \mathcal N(0, \E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)])$.
	Now, by the ergodic theorem 
	\begin{equation}
	\lim_{T \to \infty} \widetilde M^{-1} = \mathcal M_\epl^{-1}, \quad \text{a.s.}
	\end{equation}
	Therefore, we have by Slutsky's theorem that
	\begin{equation}
	\lim_{T \to \infty} \sqrt{T} I_2^\epl(T) = \Lambda^\epl \sim \mathcal N\left(0, \Gamma^\epl\right), \quad \text{in law}.
	\end{equation}
	where the covariance matrix $\Gamma^\epl$ is given by
	\begin{equation}
		\Gamma^\epl = 2\sigma \mathcal M_\epl^{-1}\E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)]\mathcal M_\epl^{-\top}.
	\end{equation}
	Finally, we denote by $\Lambda$ the limit for $\epl \to 0$ of the sequence of Gaussian random variables $\Lambda^\epl$.

	\corr{Idea for the first term:} Let us introduce the notation
	\begin{equation}
		J_1^\epl(T) \defeq \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl\right),
	\end{equation}
	and let us remark that we can rewrite
	\begin{equation} \label{eq:decomposition_first_term}
	\begin{aligned}
	J_1^\epl(T) &= \sqrt{T} \left( I_1^\epl(T) - \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\mathfrak p_\epl \right) \\
	& \quad + \sqrt{T} \left(\left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} - \mathcal M_\epl^{-1}\right) \mathfrak p_\epl\\
	&\eqdef J_{1,1}^\epl(T) + J_{1,2}^\epl(T).
	\end{aligned}
	\end{equation}
	In order to treat the terms $J_{1,1}^\epl(T)$ and $J_{1,2}^\epl(T)$, we will employ a technique presented in \cite[Remark 6.17]{PaS08}. Denoting by $\mathcal L^\epl$ be the generator of the joint process $(X^\epl, Z^\epl)$, i.e.,
	\begin{equation}
		\mathcal L^\epl = -\left(V'(x) + \frac1\epl p'\left(\frac{x}{\epl}\right)\right)\partial_x + \frac1\delta (x - z) \partial_z  + \sigma \partial^2_{xx},
	\end{equation}
	let $\Phi^\epl$ the solution of the Poisson equation
	\begin{equation}\label{eq:CLT_Poisson}
		\mathcal L^\epl \Phi^\epl(x, z) = \chi(x, z) - \E^{\rho^\epl}[\chi(X^\epl, Z^\epl)].
	\end{equation}	
	where $\chi$ is a sufficiently smooth function. Then, an application of Itô formula  yields
	\begin{equation}
	\frac{1}{T} \int_0^T  \chi(X_t^\epl, Z_t^\epl) \dd t = \E^{\rho^\epl}[\chi(X^\epl, Z^\epl)] + \frac{R_T^\epl}{T} - \sqrt{2\sigma}\frac{S^\epl_T}{T},
	\end{equation}
	where
	\begin{equation}
		R_T^\epl \defeq \Phi^\epl(X^\epl_T, Z^\epl_T) - \Phi^\epl(X^\epl_0, Z^\epl_0), \quad S^\epl_T \defeq \int_0^T \partial_x \Phi^\epl(X_t^\epl, Z_t^\epl) \dd W_t.
	\end{equation}
	Let us consider $J_{1,1}^\epl(T)$ and denote by $\Phi_1^\epl$ the solution of \eqref{eq:CLT_Poisson} with 
	\begin{equation}
		\chi(x, z) = 
	\end{equation}

	Let us first focus on $J_{1,2}^\epl(T)$. Denoting by

	Therefore, we have
%	\begin{equation}
%	\begin{aligned}
%		\sqrt{T}\left(\alpha - A + I_1^\epl(T)\right) = & \sqrt{T}\left(\alpha - A + \widetilde M^{-1}\E^{\rho^\epl}\left[V'(Z^\epl)\frac1\epl p'\left(\frac{X^\epl}{\epl}\right)\right]\right)\\
%		& + \frac{\widetilde M^{-1} R_T^\epl}{\sqrt{T}} - \sqrt{2\sigma} \frac{\widetilde M^{-1} S_T^\epl}{\sqrt{T}}.
%	\end{aligned}
%	\end{equation}
	\begin{equation}
		J_{1,1}^\epl(T)= \sqrt{T} \mathcal M_\epl^{-1} \left( \frac{R_T^\epl}{T} - \sqrt{2\sigma}\frac{S^\epl_T}{T} \right)
	\end{equation}
	Since \corr{$R^\epl_T$ is bounded independently of $\epl$}, we first get
%	\begin{equation}
%		\lim_{T \to \infty} \frac{\widetilde M^{-1} R_T^\epl}{\sqrt{T}} = 0, \quad \text{a.s.}
%	\end{equation}
	\begin{equation}
		\lim_{T \to \infty} \frac{\mathcal M_\epl^{-1} R_T^\epl}{\sqrt{T}} = 0, \quad \text{a.s.}
	\end{equation}
	Repeating the same reasoning as for $Q_T^\epl$, we get 
%	\begin{equation}
%		\lim_{T \to \infty} \sqrt{2\sigma} \frac{\widetilde M^{-1} S_T^\epl}{\sqrt{T}} = \Lambda_1^\epl \sim \mathcal N(0, \Gamma_1^\epl),
%	\end{equation}
	\begin{equation}
	\lim_{T \to \infty} \sqrt{2\sigma} \frac{\mathcal M_\epl^{-1} S_T^\epl}{\sqrt{T}} = \Lambda_1^\epl \sim \mathcal N(0, \Gamma_1^\epl),
	\end{equation}
	where the covariance is given by
	\begin{equation}
		\Gamma_1^\epl = 2\sigma \mathcal M^{-1}\E^{\rho^\epl}[\partial_x \Phi_1^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi_1^\epl(X^\epl, Z^\epl)]\mathcal M^{-\top}.
	\end{equation}
	\corr{For the second part of \eqref{eq:decomposition_first_term} we follow a similar procedure defining a different Poisson equation. The argument has to be slightly modified because we have to work with random matrix rather than random vector, but finally we obtain a second Gaussian term $\mathcal N(0, \Gamma_2^\epl)$. Therefore, we have to study the limit as $\epl$ goes to zero of the covariance matrices $\Gamma_1^\epl$ and $\Gamma_2^\epl$, which depend on the solutions $\Phi_1^\epl$ and $\Phi_2^\epl$ of the Poisson problems.} \\

	\corr{Idea for the second term:} The second term can be decomposed in 
	\begin{equation}
	\begin{aligned}
	\sqrt{T} \frac1\delta & \left( \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ]^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] \right. \\
	&\hspace{3cm} \left. - \E^{\rho^0} [ V'(Z) \otimes V'(X) ]^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)] \right) \\
	&= \sqrt{T} \frac1\delta \left( \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ]^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] \right. \\
	&\hspace{3cm} \left. - \E^{\rho^0} [ V'(Z) \otimes V'(X) ]^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] \right) \\
	&\quad + \sqrt{T} \frac1\delta \left( \E^{\rho^0} [ V'(Z) \otimes V'(X) ]^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] \right. \\
	&\hspace{3cm} \left. - \E^{\rho^0} [ V'(Z) \otimes V'(X) ]^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)] \right), \\
	\end{aligned}
	\end{equation}
	\corr{and we need the rate of convergence of the measure $\rho^\epl$ to $\rho^0$.}
\end{proof}

\section{The Bayesian setting}\label{sec:Bayesian}

@@

\subsection{Parameters of the filter}
\begin{figure}[t]
	\centering
	\begin{tabular}{ccc}
		\includegraphics[]{Figures/OUSubs_s5} & \includegraphics[]{Figures/OUFilt_s5_b1}  & \includegraphics[]{Figures/OUFilt_s5_b5} \\ % & \includegraphics[]{Figures/OUFilt_s5_b10} \\
		\includegraphics[]{Figures/OUSubs_s7} & \includegraphics[]{Figures/OUFilt_s7_b1}  & \includegraphics[]{Figures/OUFilt_s7_b5} \\ % & \includegraphics[]{Figures/OUFilt_s7_b10} \\
		\includegraphics[]{Figures/OUSubs_s10} & \includegraphics[]{Figures/OUFilt_s10_b1}  & \includegraphics[]{Figures/OUFilt_s10_b5} % & \includegraphics[]{Figures/OUFilt_s10_b10}
	\end{tabular}
	\caption{Results for the Ornstein--Uhlenbeck process. The axis of abscissae $\zeta$ represents the value of $\delta = \epl^\zeta$ for both subsampling and the filter. The three rows correspond to $\sigma = 0.5, 0.7, 1.0$ from top to bottom.}
	\label{fig:OU}
\end{figure}
We consider $N = 1$ and the quadratic potential $V(x) = x^2/2$. In this case, the solution of the homogenized equation is an Ornstein--Uhlenbeck process. Moreover, we set the multiscale equation \eqref{eq:SDE_MS} with $\epl = 0.1$ and the fast potential $p(y) = \cos(y)$. We generate data $X^\epl$ for $0 \leq t \leq T$ and $T = 10^3$ employing the Euler--Maruyama method with time step $\Delta_t = \epl^3$.

@@

@@

\subsection{Central limit theorem}
\begin{figure}[t]
	\centering
	\includegraphics[]{Figures/CLT}
	\caption{Central limit theorem result. The histogram represents numerical results, the solid curve a Gaussian fit to the latter and the dashed curve $\rho_a$ represents the theoretical estimate given in Theorem \ref{thm:CLT}.}
	\label{fig:CLT}
\end{figure}

@@

\begin{proof}[Proof of Lemma \ref{lem:density}] We have to show that the joint process solution to \eqref{eq:systemSDE} is hypo-elliptic. Denoting as $f\colon \R \to \R$ the function 
	\begin{equation}
	f(x) = -\alpha \cdot V'(x) - \frac1\epl p'\left(\frac{x}\epl\right),
	\end{equation}
	the generator of the process $(X^\epl, Z^\epl)^\top$ is given by
	\begin{equation}
	\mathcal L = f \partial_x + \sigma \partial_{xx}^2 + \frac1\delta (x - z)\partial_z \eqdef \mathcal X_0 + \sigma \mathcal X_1^2, 
	\end{equation}
	where 
	\begin{equation}
	\mathcal X_0 = f \partial_x + \frac1\delta (x - z)\partial_z, \quad \mathcal X_1 = \partial_x.
	\end{equation}
	The commutator $[\mathcal X_0, \mathcal X_1]$ applied to a test function $v$ then gives
	\begin{equation}
	\begin{aligned}
	[\mathcal X_0, \mathcal X_1]v &= f \partial_x^2 v + \frac1\delta (x - z) \partial_x \partial_z v  - \partial_x\left(f\partial_x v + \frac1\delta (x - z)\partial_z v\right)\\
	&= -\partial_x f \partial_x v - \frac1\delta \partial_z v.
	\end{aligned}
	\end{equation}
	Consequently, 
	\begin{equation}
	\mathrm{Lie}\left(\mathcal X_1, [\mathcal X_0, \mathcal X_1]\right) = \mathrm{Lie} \left(\partial_x, -\partial_x f \partial_x - \frac1\delta\partial_z\right),
	\end{equation}
	which spans $T_x\R^2$. The desired result then follows from Hörmander's theorem (see e.g. \cite[Chapter 6]{Pav14}).
\end{proof}

@@

>>>>

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{ABT10,ABJ13,PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenized equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenized equation, rather than to those of the homogenized one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}.

@@

\subsection{Central limit theorem}

\corr{Text to be added with motivation, proof in Appendix \ref{ap:CLT}.
\begin{theorem}\label{thm:CLT} Central limit theorem statement:
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \sqrt{T}\left(\widehat A^\epl_k(T) - A\right) = \Lambda, \quad \text{in law,}
	\end{equation}
	where $\Lambda \sim \mathcal N(0, \Gamma)$ and
	\begin{equation}
		\Gamma = 2\sigma \E^{\rho^0}[V'(Z) \otimes V'(X)]^{-1}\E^{\rho^0}[V'(Z) \otimes V'(Z)]\E^{\rho^0}[V'(X) \otimes V'(Z)]^{-1}.
	\end{equation}
	This covariance may not be correct (see proof in the Appendix) but it is just the limit of the ``third term'' 
\end{theorem}
}
\section{The Bayesian setting}\label{sec:Bayesian}

@@

\subsection{Parameters of the filter}\label{sec:Num_Param}
\begin{figure}[t]
	\centering
	\begin{tabular}{ccc}
		\includegraphics[]{Figures/OUSubs_s5} & \includegraphics[]{Figures/OUFilt_s5_b1}  & \includegraphics[]{Figures/OUFilt_s5_b5} \\ % & \includegraphics[]{Figures/OUFilt_s5_b10} \\
		\includegraphics[]{Figures/OUSubs_s7} & \includegraphics[]{Figures/OUFilt_s7_b1}  & \includegraphics[]{Figures/OUFilt_s7_b5} \\ % & \includegraphics[]{Figures/OUFilt_s7_b10} \\
		\includegraphics[]{Figures/OUSubs_s10} & \includegraphics[]{Figures/OUFilt_s10_b1}  & \includegraphics[]{Figures/OUFilt_s10_b5} % & \includegraphics[]{Figures/OUFilt_s10_b10}
	\end{tabular}
	\caption{Results for the Ornstein--Uhlenbeck process. The axis of abscissae $\zeta$ represents the value of $\delta = \epl^\zeta$ for both subsampling and the filter. The three rows correspond to $\sigma = 0.5, 0.7, 1.0$ from top to bottom.}
	\label{fig:OU}
\end{figure}
We consider $N = 1$ and the quadratic potential $V(x) = x^2/2$. In this case, the solution of the homogenized equation is an Ornstein--Uhlenbeck process. Moreover, we set the multiscale equation \eqref{eq:SDE_MS} with $\epl = 0.1$ and the fast potential $p(y) = \cos(y)$. We generate data $X^\epl$ for $0 \leq t \leq T$ and $T = 10^3$ employing the Euler--Maruyama method with time step $\Delta_t = \epl^3$.

@@

\subsection{Central limit theorem}
\begin{figure}[t]
	\centering
	\includegraphics[]{Figures/CLT}
	\caption{Central limit theorem result. The histogram represents numerical results, the solid curve a Gaussian fit to the latter and the dashed curve the theoretical estimate given in Theorem \ref{thm:CLT}.}
	\label{fig:CLT}
\end{figure}

In this experiment we wish to confirm the validity of Theorem \ref{thm:CLT}. We consider the same test equation as for Section \ref{sec:Num_Param}, i.e., the quadratic potential $V(x) = x^2/2$ with fluctuating potential $p(y) = \sin(y)$, multiscale parameter $\epl = 0.05$ and diffusion coefficient $\sigma = 1$. The parameters of the filter are set to $\beta = 1$ and $\delta = 1$. We compute the estimator $\widehat A^\epl_k(T)$ with final time $T = 10^3$ on $2000$ realizations of the solution and estimate the quantity 
\begin{equation}	
	\Delta_A^\epl(T) \defeq \sqrt{T}\left(\widehat A^\epl_k(T) - A\right),	
\end{equation}
where $A$ is the drift coefficient of the homogenized equation. Results, depicted in Figure \ref{fig:CLT}, show that the distribution of $\Delta_A^\epl(T)$ indeed follows a zero-mean Gaussian law, \corr{whose covariance agrees with the theoretical results (?)}.

@@

@@

\begin{proof}[Proof of Lemma \ref{lem:density}] We have to show that the joint process solution to \eqref{eq:systemSDE} is hypo-elliptic. Denoting as $f\colon \R \to \R$ the function 
	\begin{equation}
	f(x) = -\alpha \cdot V'(x) - \frac1\epl p'\left(\frac{x}\epl\right),
	\end{equation}
	the generator of the process $(X^\epl, Z^\epl)^\top$ is given by
	\begin{equation}
	\mathcal L = f \partial_x + \sigma \partial_{xx}^2 + \frac1\delta (x - z)\partial_z \eqdef \mathcal X_0 + \sigma \mathcal X_1^2, 
	\end{equation}
	where 
	\begin{equation}
	\mathcal X_0 = f \partial_x + \frac1\delta (x - z)\partial_z, \quad \mathcal X_1 = \partial_x.
	\end{equation}
	The commutator $[\mathcal X_0, \mathcal X_1]$ applied to a test function $v$ then gives
	\begin{equation}
	\begin{aligned}
	[\mathcal X_0, \mathcal X_1]v &= f \partial_x^2 v + \frac1\delta (x - z) \partial_x \partial_z v  - \partial_x\left(f\partial_x v + \frac1\delta (x - z)\partial_z v\right)\\
	&= -\partial_x f \partial_x v - \frac1\delta \partial_z v.
	\end{aligned}
	\end{equation}
	Consequently, 
	\begin{equation}
	\mathrm{Lie}\left(\mathcal X_1, [\mathcal X_0, \mathcal X_1]\right) = \mathrm{Lie} \left(\partial_x, -\partial_x f \partial_x - \frac1\delta\partial_z\right),
	\end{equation}
	which spans the tangent space of $\R^2$ at $(x, z)$, denoted $T_{x, z}\R^2$. The desired result then follows from Hörmander's theorem (see e.g. \cite[Chapter 6]{Pav14}).
\end{proof}

@@

\section{Proof of Theorem \ref{thm:CLT}}\label{ap:CLT}

Let us first introduce a technical Lemma.
\begin{lemma}\label{lem:CLT} Let $\mathcal L_\epl$ be the generator of the couple $(X^\epl, Z^\epl)^\top$, i.e.,
	\begin{equation}
		\mathcal L_\epl = -\left(\alpha \cdot V'(x) + \frac1\epl p'\left(\frac{x}{\epl}\right)\right)\partial_x + \frac1\delta (x - z) \partial_z  + \sigma \partial^2_{xx}.
	\end{equation} 
	Moreover, let $\rho^\epl$ be the invariant measure of $(X^\epl, Z^\epl)^\top$ and $\Phi^\epl\colon \R^2 \to \R^N$ be the solution of 
	\begin{equation}\label{eq:CLT_Poisson}
		-\mathcal L_\epl \Phi^\epl = \chi -  \E^{\rho^\epl}[\chi],
	\end{equation}
	for $\chi \colon \R^2 \to \R^N$ satisfying $\E^{\rho^\epl}[\Phi^\epl] = 0$. Then, it holds
	\begin{equation}\label{eq:CLT_PoissonIto}
		\frac{1}{T} \int_0^T  \chi(X_t^\epl, Z_t^\epl) \dd t = \E^{\rho^\epl}[\chi(X^\epl, Z^\epl)] - \frac{R^\epl(T)}{T} + \sqrt{2\sigma}\frac{S^\epl(T)}{T},
	\end{equation}
	where
	\begin{equation}\label{eq:CLT_PoissonRemainders}
		R^\epl(T) \defeq \Phi^\epl(X^\epl_T, Z^\epl_T) - \Phi^\epl(X^\epl_0, Z^\epl_0), \quad S^\epl(T) \defeq \int_0^T \partial_x \Phi^\epl(X_t^\epl, Z_t^\epl) \dd W_t.
	\end{equation}
	Moreover, it holds
	\begin{equation}\label{eq:CLT_Equiv}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl] = \E^{\rho^\epl}[\chi \otimes \Phi^\epl + \Phi^\epl \otimes \chi].
	\end{equation}	
\end{lemma}
\begin{proof} The proof of \eqref{eq:CLT_PoissonIto} and \eqref{eq:CLT_PoissonRemainders} is an application of the Itô formula (see e.g. \cite[Remark 6.17]{PaS08}). For \eqref{eq:CLT_Equiv}, it is possible to show that since $\mathcal L_\epl^*\rho^\epl = 0$ it holds 
	\begin{equation}
		\mathcal L_\epl^* (\Phi^\epl \rho) = 2 \sigma \rho^\epl \partial_{xx}^2 \Phi^\epl - \rho^\epl \mathcal L_\epl\Phi^\epl + 2 \sigma \partial_x \Phi^\epl \partial_x \rho^\epl.
	\end{equation}
	Therefore, an integration by parts yields 
	\begin{equation}
	\begin{aligned}
		\E^{\rho^\epl}[\mathcal L_\epl \Phi^\epl \otimes \Phi^\epl] &= 	\int_{\R}\int_{\R} \Phi^\epl \otimes \mathcal L_\epl^*\left(\Phi^\epl \rho^\epl\right) \dd x \dd z \\
		&= -\int_{\R}\int_{\R}\Phi^\epl \otimes \mathcal L_\epl \Phi^\epl \rho^\epl \dd x \dd z- 2\sigma \int_{\R}\int_{\R} \partial_x \Phi^\epl \otimes \partial_x \Phi^\epl \rho^\epl \dd x \dd z\\
		&= -\E^{\rho^\epl}[\Phi^\epl \otimes \mathcal L_\epl \Phi^\epl] - 2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl]. 
	\end{aligned}
	\end{equation}
	Finally, since $\E^{\rho^\epl}[\Phi] = 0$ 
	\begin{equation}
	\begin{aligned}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl] &= -\E^{\rho^\epl}[\mathcal L_\epl \Phi^\epl \otimes \Phi^\epl + \Phi^\epl \otimes \mathcal L_\epl \Phi^\epl]\\
		&= \E^{\rho^\epl}[\chi \otimes \Phi^\epl + \Phi^\epl \otimes \chi],
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

We can now proceed with the main proof.
\begin{proof}[Proof of Theorem \ref{thm:CLT}] Let us introduce the notation
	\begin{equation}\label{eq:CLT_DefMEps}
	\mathcal M_\epl \defeq \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ], \quad \mathcal M_0 \defeq \E^{\rho^0} [ V'(Z) \otimes V'(X) ],
	\end{equation}
	and the notation
	\begin{equation}
	\mathfrak p_\epl \defeq \E^{\rho^\epl} \left [ \frac{1}{\epl} p' \left ( \frac{X^\epl}{\epl} \right ) V'(Z^\epl) \right ]\
	\end{equation}
	and let us remark that from the proof of Theorem \ref{thm:mainTheorem} one can deduce the equalities
	\begin{equation}\label{eq:CTL_A}
	A = \frac1\delta \mathcal M_0^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)],
	\end{equation}
	and
	\begin{equation}\label{eq:CTL_alpha}
	\alpha = \mathcal M_\epl^{-1} \left(\frac1\delta \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] - \mathfrak p_\epl \right).
	\end{equation}
	Let us furthermore remark that the decomposition \eqref{eq:alphaDecomposition} yields
	\begin{equation}
	\sqrt{T}\left(\widehat A^\epl_k(T) - A\right) = \sqrt{T} \left(\alpha - A + I_1^\epl - I_2^\epl\right).
	\end{equation}
	Replacing the expression for $A$ and $\alpha$ given in \eqref{eq:CTL_A} and \eqref{eq:CTL_alpha}, we get
	\begin{equation}
	\begin{aligned}
	\sqrt{T}\left(\widehat A^\epl_k(T) - A\right) &= \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl \right) \\
	&+  \frac{\sqrt{T}}\delta \left( \mathcal M_\epl^{-1} \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)] - \mathcal M_0^{-1} \E^{\rho^0}[(X^2 - Z^2)V''(Z)] \right) \\
	&- \sqrt{T} I_2^\epl(T).
	\end{aligned}
	\end{equation}
	\corr{Third term is OK:} We rewrite the term involving $I_2^\epl(T)$ as
	\begin{equation}
	\sqrt{T} I_2^\epl(T) = \frac{\sqrt{2\sigma}}{\sqrt{T}} \widetilde M^{-1} Q^\epl(T),
	\end{equation}
	where, since $Z_t^\epl$ is adapted with respect to the natural filtration $\mathcal F_t$ of the Wiener process $W \defeq \left(W_t, t \geq 0\right)$, the quantity
	\begin{equation}
	Q^\epl(T) \defeq \int_0^T V'(Z^\epl_t) \dd W_t,
	\end{equation}
	is a martingale whose quadratic variation is given by
	\begin{equation}
	\langle Q^\epl \rangle_T = \int_0^T V'(Z^\epl_t) \otimes V'(Z^\epl_t) \dd t. 
	\end{equation}
	Since the ergodic theorem guarantees that 
	\begin{equation}
	\lim_{T\to\infty}\frac{\langle Q^\epl \rangle_T}{T} = \E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)], \quad \text{in } L^1(\mu^\epl),
	\end{equation}
	the martingale central limit theorem gives 
	\begin{equation}
	\lim_{T\to\infty}\frac{1}{\sqrt{T}} Q^\epl(T) = \Xi^\epl, \quad \text{in law},
	\end{equation}
	where $\Xi^\epl \sim \mathcal N(0, \E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)])$.
	Now, by the ergodic theorem 
	\begin{equation}
	\lim_{T \to \infty} \widetilde M^{-1} = \mathcal M_\epl^{-1}, \quad \text{a.s.}
	\end{equation}
	Therefore, we have by Slutsky's theorem that
	\begin{equation}
	\lim_{T \to \infty} \sqrt{T} I_2^\epl(T) = \Lambda^\epl \sim \mathcal N\left(0, \Gamma^\epl\right), \quad \text{in law}.
	\end{equation}
	where the covariance matrix $\Gamma^\epl$ is given by
	\begin{equation}
	\Gamma^\epl = 2\sigma \mathcal M_\epl^{-1}\E^{\rho^\epl}[V'(Z^\epl) \otimes V'(Z^\epl)]\mathcal M_\epl^{-\top}.
	\end{equation}
	Finally, we denote by $\Lambda$ the limit for $\epl \to 0$ of the sequence of Gaussian random variables $\Lambda^\epl$.

	\corr{Idea for the first term:} Let us introduce the notation
	J_1^\epl(T) \defeq \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl\right),
	and let us remark that we can rewrite
	\begin{equation} \label{eq:decomposition_first_term}
	J_1^\epl(T) &= \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\left(\frac1T \int_0^T \frac{1}{\epl} p' \left ( \frac{X^\epl_t}{\epl} \right ) V'(Z^\epl_t) \dd t - \mathfrak p_\epl \right) \\
	& \quad + \sqrt{T} \left(\left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} - \mathcal M_\epl^{-1}\right) \mathfrak p_\epl\\
	&\eqdef J_{1,1}^\epl(T) + J_{1,2}^\epl(T).
	Let us denote by $\Phi_1^\epl$ the solution of \eqref{eq:CLT_Poisson} in Lemma \ref{lem:CLT} with right hand side
	\chi_1(x, z) = \frac 1 \epl p' \left ( \frac x \epl \right ) V'(z).
	Then, denoting as $R_1^\epl(T)$ and $S_1^\epl(T)$ the quantities introduced in \eqref{eq:CLT_PoissonRemainders} with $\Phi_1^\epl$ in place of $\Phi^\epl$, we have
	J_{1,1}^\epl(T)= \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \left( -\frac{R_1^\epl(T)}{T} + \sqrt{2\sigma}\frac{S_1^\epl(T)}{T} \right).
	Since \corr{$R_1^\epl(T)$ is bounded independently of $\epl$ (is it clear?)}, we first get by the ergodic theorem
	\lim_{T \to \infty} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \frac{R_1^\epl(T)}{\sqrt{T}} = 0, \quad \text{a.s.}
	Repeating the same reasoning as for $Q^\epl(T)$ and employing the ergodic theorem, we get 
